---
title: "Effect of Sea Star Wasting Disease on Mussel Abundance"
output: html_document


---
Loading the necessary libraries
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
```


Reading in the sea star and mussel data set

```{r, results=FALSE}
abundance_data <- read.csv("data/change in pisaster and mussels PISCO sites.csv")
```

Viewing the entire dataset to get an idea about what it looks like

```{r, results=FALSE}
View(abundance_data)
```


Using ncol and nrow to find out how many variables and values there are

```{r}
ncol(abundance_data)
```

```{r}
nrow(abundance_data)
```

Using glimpse() to get a better view of the columns and their names

```{r}
glimpse(abundance_data)
```
After viewing the data, I have decided I want to manipulate and plot the data of the mean percent mussel cover before and after mass starfish mortality. I feel that this will give me a good idea about how the removal of Pisaster has affected the mussel abundance


In the below code I am renaming the columns to pre_mean_cover and post_mean_cover to make them easier to use in code

```{r, results=FALSE}
abundance_data_v2 <- rename(abundance_data, pre_mean_mussel_cover = Mean.percent_cover_Mytcal..pre.SSWS,post_mean_mussel_cover = Mean.percent_cover_Mytcal..post.SSWS)

```

I am then isolating the two columns I will be using in the graph into their own data frame in order to manipulate them further
```{r, results=FALSE}
mean_mussel_cover_data <- select(abundance_data_v2, c(pre_mean_mussel_cover, post_mean_mussel_cover))

```

In order to create a histogram with the pre and post data overlapping each other I create separate data frames: one consisting of just the pre SSWD percent mussel cover data and on with the post SSWD data. Then I rename the columns in these new, separate data frames to match so I can graph them together. I use the view() function after every change or rename to make sure the code worked properly.

```{r, results=FALSE}
pre_cover_data <- select(abundance_data_v2, c(pre_mean_mussel_cover))

pre_cover_data_v2 <- rename(pre_cover_data, mean_percent_cover = pre_mean_mussel_cover)
```

```{r, results=FALSE}
post_cover_data <-  select(abundance_data_v2, c(post_mean_mussel_cover))

post_cover_data_v2 <- rename(post_cover_data, mean_percent_cover = post_mean_mussel_cover)
```

Finally, I graph the two data frames (pre and post) in a single histogram, making sure they are transparent and a legend is included.

```{r, message=FALSE}
mean_cover_histogram <- rbind(pre_cover_data_v2, post_cover_data_v2)

mean_cover_histogram<- ggplot(mean_cover_histogram) + 
  geom_histogram(data=pre_cover_data_v2, aes(x=mean_percent_cover, fill = "Pre"), alpha = 0.5) +
  geom_histogram(data=post_cover_data_v2, aes(x=mean_percent_cover, fill = "Post"), alpha = 0.5) + 
  xlab("Mean Percent Cover") + 
  scale_fill_manual("Type", values = c(Pre = "blue", Post = "green"))

print(mean_cover_histogram)
  
```








